Modeling Second-Language Learning from a Psychological Perspective

### 2018-03-27 doc/naaclhlt2018-latex/paper.tex
add description of lesions

The results of the lesion experiments are shown in Figure~\ref{fig:lesions}. The
models are as follows.
\begin{itemize}
  \item \emph{none}: all features are included.  
  \item \emph{user}: all user-level features, including the user ID and other
    calculated features like entropy and measures of exercise bursts, are
    removed.
  \item \emph{userid} and \emph{user other}: only user ID or only the calculated
    features user features, respectively, are removed.
  \item \emph{word}: Token and token root IDs; previous, next, and dependency-tree
    root word IDs; and morphological, part of speech, and dependency tree
    features are removed.
  \item \emph{word id} and \emph{word other}: only word IDs or only other
    word features, respectively, are removed.
  \item \emph{neighbors}: Both word IDs and other word features are removed, but
    only for the previous, next, and dependency-tree root words. Information
    about the present word is maintained.
  \item \emph{external}: External information about the word, including corpus
    frequency, Levenshtein distance from translation, and age of acquisition,
    are removed.
  \item \emph{temporal}: Temporal information, including number and timing of
    past encounters with the word and error tracking information, is removed.
\end{itemize}

>>>>

The results of the lesion experiments are shown in Figure~\ref{fig:lesions}. The
models are as follows.
\begin{description}
  \item[none:] all features are included.  
  \item[user:] all user-level features, including the user ID and other
    calculated features like entropy and measures of exercise bursts, are
    removed.
  \item[userid \& user other:] only user ID or only the calculated
    features user features, respectively, are removed.
  \item[word:] Token and token root IDs; previous, next, and dependency-tree
    root word IDs; and morphological, part of speech, and dependency tree
    features are removed.
  \item[word id \& word other:] only word IDs or only other
    word features, respectively, are removed.
  \item[neighbors:] Both word IDs and other word features are removed, but
    only for the previous, next, and dependency-tree root words. Information
    about the present word is maintained.
  \item[external:] External information about the word, including corpus
    frequency, Levenshtein distance from translation, and age of acquisition,
    are removed.
  \item[temporal:] Temporal information, including number and timing of
    past encounters with the word and error tracking information, is removed.
\end{description}

Interestingly, we found that for both user-level and word-level features, the
bulk of the model's predictive power could be achieved using ID's alone,
represented as high-cardinality categorical features. Removing other word features,
such as morphological features and part of speech, created only a small
degradation of performance. In the case of users, removing features such as
entropy and average exercise burst length led to a tiny increase of performance.
In the case of both users and words, though, we find that in the absence of ID
features the other features are helpful and leads to better performance than
removing all features. We also found that removing all information about
neighboring words and the dependency-parse root word degraded performance. This
confirms that word context matters, and likely points to errors in word order,
subject--verb matching and other grammatical rules.

Our external word features---Levenshtein distance to translation, frequency, and
age of acquisition---provided a slight boost to model performance, showing the
benefit of considering from a psychological and linguistic perspective what
makes a word hard to learn. Adding temporal features about past encounters and
errors helped the models, but not as much as we expected. While not included in
the final model, we had also tried augmenting the temporal feature set with more
features related to massing and spacing of encounters with a word, but found it
did not improve performance. This is perhaps not surprising given how small the
benefit of the existing temporal features are in our model.

While not plotted above, we also ran a model lesioning exercise-level features
including client, session type, format, and exercise duration. This model
achieved an AUROC of $.787$, far lower than any other lesion. This points to the
importance of how a question is asked for user performance, reflecting insights
from psychology such as the difference between recall and recognition memory
\cite{yonelinas2002nature}.
